---
title: "Class Notes t-tests"
author: "Holly Westbrook"
date: "3/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(palmerpenguins)
library(rstatix)
```
########## t-test ########## 
one sample t test: compare distribution to a theoretical mean

Paired sample t test: compare 2 distributions to each other (also called independent t-test)

```{r}
### One sample t-test
# gentoo

gentoo = penguins %>% 
  filter(species == "Gentoo")

ggplot(data=gentoo)+
  geom_histogram(aes(x=body_mass_g))
#distribution looks normal

summary(gentoo)

## t-test for gentoo body mass
#start by doing some data exploration before you try to run statistics
mean(gentoo$body_mass_g, na.rm = TRUE)

sd(gentoo$body_mass_g, na.rm = TRUE)

```

t-test assumptions:
1. no outliers
  if there are, why are they there
2. Check for normal distribution

```{r}
# test for ouliers
gentoo %>% 
    identify_outliers(body_mass_g)

#no outliers found!
# you can also look at the histogram and see if anything is sticking plot
```
Q-Q plots are also used, they line up data and plot them against a normal distribution
It should be a straight line, it may get weird towards the end but as long as it's not too dramatic it's fine

```{r}
# Q-Q plot
ggplot(data=gentoo)+
  stat_qq(aes(sample=body_mass_g))

```

Now we'll do a one sample t-test
We'll take the average body mass from the encyclopedia of life

```{r}
t.test(gentoo$body_mass_g, mu = 5500)
# mu is the theoretical mean, 5500 is from the encyclopedia of life
```
Calculates:
  the mean
  p-value (low p-value indicates significance)

You can also do this in dplyr

```{r}
gentoo %>% 
  t_test(body_mass_g~1, mu = 5500) #~1 indicates 1 sample t-test

#you can save this to a variable, and then make a data frame if you need to repeat this function

gentoo_results = gentoo %>% 
  t_test(body_mass_g~1, mu = 5500)

```

######### Unpaired 2 sample t-test aka "independent sample t-test ##############

```{r}
# gentoo vs adelie body mass
data_for_t_test = penguins %>% 
  filter(species %in% c("Gentoo", "Adelie"),
         !is.na(body_mass_g)) %>% #filters out NAs
  select(species, body_mass_g) %>%  #gets rid of data columns we won't be using
  droplevels() #for any factor variable in the data set it will get rid of any level that has zero rows in it
#levels are like different subcategories in a column. So for species our original levels were Adelie, Gentoo, and Chinstrap  
```

Now we calculate summary statistics 
```{r}
data_for_t_test %>% 
  summarise(mean = mean(body_mass_g),
            sd = sd(body_mass_g))

#but we want it separate for each species so:

data_for_t_test %>%
  group_by(species) %>% #groups by species. shocking. i know
  summarise(mean = mean(body_mass_g),
            sd = sd(body_mass_g))

ggplot(data = data_for_t_test)+
  geom_histogram(aes(x=body_mass_g))+
  facet_wrap(~species) #~ indicates you're usinga formula, facet wrap will create separate panels based on what you "wrap" it around. So in this case we're wraping it around the species level
  #Facet wrap will automatically make the scales on the panels the same by default, we can rescale it like so

ggplot(data = data_for_t_test)+
  geom_histogram(aes(x=body_mass_g))+
  facet_wrap(~species, scales = "free")


```

Identify outliers now and make Q-Q plots


```{r}
data_for_t_test %>% 
  group_by(species) %>% 
  identify_outliers(body_mass_g)


ggplot(data = data_for_t_test)+
  stat_qq(aes(sample = body_mass_g))
#tails are a bit like an S, but let's separate the species 

ggplot(data = data_for_t_test)+
  stat_qq(aes(sample = body_mass_g))+
  facet_wrap(~species, scales = "free")
#these look pretty normal now
```

Because we're comparing two distributions, we should check to see if the variances between the two datasets are similar. 
If variance is equal -> students t.test
If variance is not equal -> Welch's t.test
use a levene test for this (null value = variances equal)

```{r}
data_for_t_test %>% 
  levene_test(body_mass_g ~ species) #comparing the body mass distribution for each factor in the species column

```

This passed the levene test, bc p>0.5 which means the variances are the same

Now we can finally run our unpaired 2 sample t-test

```{r}
t.test(data_for_t_test$body_mass_g ~data_for_t_test$species)

```

p-value is super small so the Gentoo & Adelie body masses are significantly different from each other
This is a Welch's t-test, but because we know our variences are qual we can run a student's t-test instead
The default is the Welch's t-test because it assumes you're lazy and didn't check all the above shit

```{r}

t.test(data_for_t_test$body_mass_g ~data_for_t_test$species, var.equal = TRUE)

#try it in dplyr

data_for_t_test %>% 
  t_test(body_mass_g ~ species)

```

A Paired t-test is good for when you are re-runing samples after some sort of treatment (or maybe just re-analyzing data) and want to see if it changed significantly



### Exercise 3.1

Are Adelie penguin flipper lengths significantly different between males and females? Do some exploratory data analysis. Compute summary statistics and plot histograms. Then conduct an independent sample t-test. What do your results show?

```{r}
head(penguins)
 
ex3 = penguins %>% 
  filter(species == "Adelie",
         !is.na(flipper_length_mm),
         !is.na(sex)) %>% 
  select(sex, flipper_length_mm) %>%  
  droplevels()

ex3

ex3 %>% 
  group_by(sex) %>% 
  identify_outliers(flipper_length_mm)

ex3 %>%
  group_by(sex) %>% #groups by species. shocking. i know
  summarise(mean = mean(flipper_length_mm),
            sd = sd(flipper_length_mm))
ex3
ggplot(data = ex3)+
  geom_histogram(aes(x=flipper_length_mm))+
  facet_wrap(~sex, scale = "free")


ggplot(data = ex3)+
  stat_qq(aes(sample = flipper_length_mm))+
  facet_wrap(~sex, scales = "free")
#this looks funky? there are outliers but because our n is high enough the central limit theory says it's okay?

ex3 %>% 
  t_test(flipper_length_mm~sex)


```


###### Correlation ######

```{r}

ggplot(data=gentoo) +
  geom_point(aes(x=bill_length_mm, y = bill_depth_mm))

# correlation

cor.test(x = gentoo$bill_length_mm, y =gentoo$bill_depth_mm) #if NAs are causing an error add in this term
cor.test(x = gentoo$bill_length_mm, y =gentoo$bill_depth_mm, use="complete.obs")

#also you can do this to just get the corelation value
cor(x = gentoo$bill_length_mm, y =gentoo$bill_depth_mm, use="complete.obs")

#easy to run correlation on multiple columns in a dataset w/ cor
head(gentoo)
#want all rows and columns 3:6
cor(gentoo[,3:6], use="complete.obs")




```

^This is a correlation matrix, which is hella neat



```{r}
#get a fuck ton of stats fast with the ggpairs function in GGally
library(GGally)
gentoo %>%  
  select(bill_length_mm, bill_depth_mm, body_mass_g, flipper_length_mm) %>% 
  ggpairs()

#gives you correlation values, histograms, and plots


#we can even look at all the species is we color map it
penguins %>%  
  select(species, bill_length_mm, bill_depth_mm, body_mass_g, flipper_length_mm) %>% 
  ggpairs(aes(color=species))


```