---
title: "Class Notes Linear Models"
author: "Holly Westbrook"
date: "3/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(palmerpenguins)
library(GGally)
library(ggiraph) #lol giraffe
library(ggiraphExtra)
library(car)

#### Correlation Matrix
head(penguins)
glimpse(penguins)
summary(penguins)
dim(penguins)

penguins %>% 
  select(-island, -sex, -year) %>% 
  ggpairs(aes(color=species))



```


Today we'll look at the relationship between bill length and bill depth

```{r}
penguins %>% 
  select(bill_length_mm, bill_depth_mm) %>% 
  ggpairs()


```





There's a lot of linear model assumptions (7), one of them is that there is a linear relationship between the two . Currently we don't see a linear relationship but we can tease that out

Residuals= difference between y and estimated/predicted y are normally distributed and homoscedastic (variance is homogenous)

No autocorrelation: more a problem in time series/spacial data, data taken close together in space (like adjacen stations) or time (like a time series with a high resolution) will likely be autocorelated

Multi-colinearity: If y is a function of x1 adn x2, if x1 and x2 are very highly colinear it makes it problematic, esp if you want to use the model to make a statement about how x1 relates to y because x2 could be stealing a lot of the power from x1. If x1 and x2 are super well correlated you don't know which proportion should be contributing mechanisticaly more to y




