---
title: "Class Notes ggplot"
author: "Holly Westbrook"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

The conflicts that pop up mean there is masking, which means that some of the funtions in the tidyverse already have the same name as functions in base R
The most recent thing that has been loaded is the version of the function that will be used
You can know which function you're using with the find function

```{r}
find("filter") #first one listed is the default 

library(palmerpenguins)

head(penguins)
#if we want to use the version of a function from a specific package you can call it like so
penguins %>% dplyr::filter(species == "Adelie")

##############################
#             Scatter Plot
##############################
ggplot() +
  geom_point(aes(x= flipper_length_mm, y = body_mass_g, color = species, size = bill_length_mm), data = penguins)
            #things inside the aes function can be mapped to data in your data frame, if you want them constant you can put them outside aes
ggplot(data = penguins) + #called the data in ggplot() so that the geom_smooth() function knows to use data from penguins
  geom_point(aes(x= flipper_length_mm, y = body_mass_g, color = species, size = bill_length_mm), data = penguins)+
  geom_smooth(aes(x=flipper_length_mm, y = body_mass_g))

#we can make this even neater looking by calling the x and y in ggplot(), this will set the default x + y for all next functions
ggplot(aes(x= flipper_length_mm, y = body_mass_g), data = penguins) + #
  geom_point(aes(color = species))+
  geom_smooth(method = "lm")+
  ylab("Body mass (g)")+
  xlab("Flipper length (mm)")+
  ggtitle("Penguins Are Cute")+
  theme_classic() #there are a lot of different themes, you can manually control all the aesthetic features but it's easy to use the premade themes

```

There are also packages that work with ggplot that you can find onnline


Exercise 2.1

Build a scatter plot of bill depth vs. bill length for Adelie penguins. Map the point colors to the island that the Adelie penguin was observed on. Add axes labels and a main title. 

```{r}
adelie = penguins %>% filter(species == "Adelie")
head(adelie)
ggplot(aes(x = bill_depth_mm, y = bill_length_mm),data= adelie)+
  geom_point(aes(color = island))+
  xlab("Bill Depth (mm)")+
  ylab("Bill Length (mm)")+
  ggtitle("Adelie Penguins")

```

Let's make a penguin time series to see how it has changed over time

```{r}
penguin_ts = penguins %>%
  group_by(species, year) %>% 
  summarize(num=n())
penguin_ts

ggplot(aes(x=year, y=num, color = species), data = penguin_ts)+
  geom_point()+
  geom_line()

```

Real Histogram Hours


```{r}
############################
#     Histograms
############################

ggplot(data=penguins) +
  geom_histogram((aes(x=flipper_length_mm, fill = species))) #use fill instead of color for histograms, color will just make an outline


#we can also make them ont stacked by setting position and alpha (transparency parameter)
ggplot(data=penguins) +
  geom_histogram((aes(x=flipper_length_mm, fill = species)), position = "identity", alpha = 0.5, binwidth=5)+
  scale_fill_manual (values=c("darkorange", "darkorchid", "darkcyan"))

```

Here are the colors you can use in r!
http://applied-r.com/r-color-tables/ 

Box plot time

```{r}
ggplot(aes(y=flipper_length_mm, x = species), data=penguins)+
  geom_boxplot()+
  geom_jitter(aes(color=species))

##################
#     Bar Plot
##################


ggplot(data=penguins)+
  geom_bar(aes(x=sex, fill = species))+
  facet_wrap(~species) #create a different panel for whatever variable you choose

my_plot = ggplot(data=penguins %>% filter(!is.na(sex)))+ #getting rid of NAs
  geom_bar(aes(x=sex, fill = species))+
  coord_flip()+ #to make bars horizontal instead of vertical, it looks bad in this
  facet_wrap(~species, ncol=1) #ncol= number of columns, this makes the plot look less bad lol
  ggsave(filename = "figures/penguin_sex_bar.png", device = "png", width=6, height = 4, dpi = 300, units = "in")
  #setting device isn't necessary bc we already added .png to the file name
my_plot


```













