---
title: "HW_3a_Westbrook"
author: "Holly Westbrook"
date: "3/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Chinstrap penguins are 4435 g 
    #encyclopedia of life
Adelie penguins are abuot 5000 g from:
    #Carol M. Vleck, David Vleck, Physiological Condition and Reproductive Consequences in Adélie Penguins, Integrative and Comparative Biology, Volume 42, Issue 1, February 2002, Pages 76–83, https://doi.org/10.1093/icb/42.1.76
    

```{r}
library(palmerpenguins)
library(ggplot2)
library(dplyr)
library(rstatix)


#creating a data frame with literature value
lit_peng = data.frame("Body_Mass"= c(4435, 5000), "Species" = c("Literature Chinstrap", "Literature Adelie"))
penguins

#creating data frames with just the adelie and chinstrap data
adelie = penguins %>% 
  filter(penguins$species == "Adelie",
         !is.na(body_mass_g))
chinstrap = penguins %>% 
  filter(penguins $species == "Chinstrap", 
         !is.na(body_mass_g))

#calculating mean body mass for the species so I can plot them against the literature values
adelie_mean_body_mass = mean(adelie$body_mass_g, na.rm = TRUE)
chinstrap_mean_body_mass = mean(chinstrap$body_mass_g, na.rm = TRUE)

#adding all the data into a data frame
mean_body_masses = data.frame("species" = c("Adelie", "Chinstrap", "Adelie Literature", "Chinstrap Literature"), 
                              "mean_body_mass"= c(adelie_mean_body_mass, chinstrap_mean_body_mass, 5000, 4435))

########## Bar Plot of Body Mass ########## 
bodymass = ggplot(data = mean_body_masses,
                  aes(x = species, y= mean_body_mass, fill = species)) +
    geom_bar(stat = "identity")+
    theme_classic()+
    xlab("Species")+
    ylab("Mean Body Mass (g)")

bodymass

########## t-tests ########## 
t.test(chinstrap$body_mass_g, mu = 4435)

t.test(adelie$body_mass_g, mu = 5000) 
```
1. Adelie body mass in the palmer penguins dataset is significantly different from the literature values as p < 0.05
2. Chinstrap body mass in the palmer penguins dataset is significantly different from the literature values as p < 0.05


```{r}
#making another dataframe because I didn't read ahead about what I'd need to do
penguin_data_for_t_test = penguins %>%
  filter(species %in% c("Chinstrap", "Adelie"),
         !is.na(body_mass_g)) %>%
  select(species, body_mass_g) %>%
  droplevels() 

#summarizing the data
penguin_data_for_t_test %>%
  group_by(species) %>%
  summarize(mean=mean(body_mass_g), sd=sd(body_mass_g))

#make a histogram and see if anything sticks out
ggplot(aes(x=body_mass_g), data=penguin_data_for_t_test) +
  geom_histogram() +
  facet_wrap(~species)

#find outliers 
penguin_data_for_t_test %>%
  group_by(species) %>%
  identify_outliers(body_mass_g)

### there are outliers so we should use a Wilcoxon test 
wilcox.test(penguin_data_for_t_test$body_mass_g ~ penguin_data_for_t_test$species)

```   

3. They do not vary significantly from each other because p > 0.5 in the Wilcoxon test

```{r}

#quickly plotting them all together just to see what I think the answer will be
ggplot(penguins, aes(x= flipper_length_mm, y = body_mass_g))+
  geom_point()

#removing NAs
all_penguin_data_for_t_test = penguins %>%
  filter(!is.na(body_mass_g),
         !is.na(flipper_length_mm)) %>%
  droplevels() 

# Checking the qq plots for both variables
ggplot(all_penguin_data_for_t_test) +
  stat_qq(aes(sample=body_mass_g))

ggplot(all_penguin_data_for_t_test) +
  stat_qq(aes(sample=flipper_length_mm))

## qq plots look okay because they're about 1:1

cor.test(x = penguins$body_mass_g, y = penguins$flipper_length_mm,
         use = "complete.obs")
```

4. There is a strong positive correlation between the penguin body mass and flipper length (0.87), and it is significant because p < 0.05.

```{r}

#looking at qq plots again but separating them by species
ggplot(all_penguin_data_for_t_test) +
  stat_qq(aes(sample=body_mass_g))+
  facet_wrap(~species)


ggplot(all_penguin_data_for_t_test) +
  stat_qq(aes(sample=flipper_length_mm))+
   facet_wrap(~species)

#both of these look ~ 1:1 so we're ok to keep going

adelie_cor = cor.test(x = adelie$body_mass_g, y = adelie$flipper_length_mm,
         use = "complete.obs")
adelie_cor

chinstrap_cor = cor.test(x = chinstrap$body_mass_g, y = chinstrap$flipper_length_mm,
         use = "complete.obs")
chinstrap_cor

#making a dataframe for gentoo
gentoo = penguins %>% 
  filter(penguins$species == "Gentoo",
         !is.na(body_mass_g),
         !is.na(flipper_length_mm))

gentoo_cor =cor.test(x = gentoo$body_mass_g, y = gentoo$flipper_length_mm,
         use = "complete.obs")
gentoo_cor

```
5. Gentoo and Chinstrap penguins both have a fairly strong positive correlation (0.703 and 0.642), and both have p < 0.05. Adelie is also positively correlated, but the correlation is weaker (0.468) than the other two species. 